Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-10-06 WSSSPE3/report/wg_appendix_software_credit.tex
Software credit appendix in progress; added some relevant pubs
Former-commit-id: d4910da9ce59d01dbc76b0bd5c003886b00bda70

\begin{itemize}
\item Alice Allen -- Astrophysics Source Code Library
\item Sou-Cheng Choi -- NORC at Univ.\ Chicago, Illinois Institute of Technology
\item James Hetherington -- University College London
\item Lorraine Hwang -- University of California, Davis
\item Daniel S.\ Katz -- University of Chicago \& Argonne National Laboratory
\item Frank Löffler -- Louisiana State University
\item Abby Cabunoc Mayes -- Mozilla Science Lab
\item Kyle E.\ Niemeyer -- Oregon State University
\item Grace Peng -- National Center for Atmospheric Research
\item Ilian Todorov -- Science \& Technology Facilities Council, UK
\end{itemize}

@@

@@

@@

@@

@@

@@

@@

>>>>

\begin{itemize}
\item Alice Allen -- Astrophysics Source Code Library
\item Sou-Cheng Choi -- NORC at University of Chicago, Illinois Institute of Technology
\item James Hetherington -- University College London
\item Lorraine Hwang -- University of California, Davis
\item Daniel S.\ Katz -- University of Chicago, Argonne National Laboratory
\item Frank Löffler -- Louisiana State University
\item Abby Cabunoc Mayes -- Mozilla Science Lab
\item Kyle E.\ Niemeyer -- Oregon State University
\item Grace Peng -- National Center for Atmospheric Research
\item Ilian Todorov -- Science \& Technology Facilities Council, UK
\end{itemize}

@@

The following section summarizes the working group's discussion based on contributions prior to the meeting~\cite{WSSSPE3-SC-github-issues} and the collaborative notes taken during the meeting~\cite{WSSSPE3-SC-google-notes}.
Please refer to the original sources for the unedited discussions.

@@

There currently is no standard/accepted mechanism for citing software or receiving credit for software (akin to publications/citations).
Software is eligible for DOI assignment, but DOI metadata fields are not well tuned for software (vs. publications).
Some software providers apply for DOIs but it is still not a widely adopted.
Also, no mechanism to cite software dependencies within software.

Obstacle: indexers (e.g., Scopus) don’t currently support software/data document types or DataCite DOIs.

Obstacle: no standard mechanism for software to cite other software.

Obstacle: for tenure, there is no standard for software products to be included in policy,
sometimes not even within one university: hard to change globally.

@@

\begin{enumerate}
\item Hold virtual meeting to determine group members responsible\slash willing to work on the following tasks, to be organized within one month of the workshop
\item Compile best practices of software citation across multiple disciplines, including journals and communities of interest\slash practice in the research world, to begin by December 2015
\item Compile examples of including other products in promotion and tenure dossier, to begin by December 2015
\item Draft the Software Citation Principles document (including citation metadata file), by April 2016
\item Publish\slash release the Software Citation Principles document, by August 2016
\item Reach out to journals, publishers, teachers\slash educators, indexers, and professional societies---likely through meetings with key groups, to begin by September 2016

\end{enumerate}

@@

The WSSSPE breakout group plans to join efforts related to citing softare with the FORCE11 Software Citation Working Group (FORCE11-SCWG)\footnote{FORCE11 Software Citation Working Group, \url{https://www.force11.org/group/software-citation-working-group}}; Kyle Niemeyer formally requested the merging of these groups following the meeting.
However, some future plans of the WSSSPE group fall outside the scope of FORCE11-SCWG, which covers software citation practices.
These activites include working with indexers such as Web of Science and Scopus to index software citations archived on, e.g., Zenodo\footnote{Zenodo, \url{https://zenodo.org}}, and pursuing the development of an open indexing service; such plans will be pursued either separately or through the formation of follow-on FORCE11 working groups.

The group will primarily communicate electronically, with Kyle Niemeyer responsible for ensuring regular progress.

@@

The near-term actions of the group, focused mainly on software citation, do not require any additional resources.
However, connections with publishers and indexers will be needed to pursue related activities, although the FORCE11-SCWG may satisfy this need; in addition, some members of the group already reached out to relevant contacts.
Funding may be needed to organize meetings or for group members to attend relevant meetings, as discussed further below.

@@

Following the meeting, Kyle Niemeyer formally requested the merging of software citation activities with FORCE11-SCWG.
Within a month of the meeting, Kyle will organize a virtual meeting of the group and lead the division of responsibilities for compiling existing practices of software citation and including software\slash products in promotion and tenure dossiers.
Building off of these efforts, the next major milestone is drafting the Software Citation Principles document in collaboration with the SCWG, targeted for April 2016.
While the existing directors of the SCWG, Arfon Smith and Dan Katz, lead the efforts of that group torwards the Software Citation Principles document, Kyle will help coordinate contributions from the WSSSPE group members.

@@

Some funding would be useful to support primarily travel to conferences for group meetings (e.g., FORCE2016\footnote{FORCE2016, \url{https://www.force11.org/meetings/force2016}}), and to hold meetings to bring together both group members and key stakeholders (e.g., journals, publishers, professional societies, indexers).
In addition, funding would be desired to support group members' time to perform work towards the key steps described previously.
